---
title: "Dashboards and Financial Data"
author: "Anjana Ananthraman"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
    source_code: embed
---

Stock Prices
=====================================

Sidebar {.sidebar}
-----------------------------------------------------------------------

### 

In this dashboard I have analyzed four companies stocks such as Workday (WDAY),SAP, Oracle (ORCL)and Tesla (TSLA).I have performed this analysis for a period of 180 days where I have considered various parameters like open, close, high, low prices.I have also studied the adjusted closing price and the trading volume for each of the stocks.

```{r setup, include=FALSE}
library(flexdashboard)
chooseCRANmirror(graphics=FALSE, ind=1)
install.packages("pdfetch")
install.packages("dygraphs")
library(pdfetch)
library(dygraphs)
library(xts)
identifiers = c("WDAY","SAP","ORCL","TSLA")
```



Column {.tabset data-height=550}
-----------------------------------------------------------------------

### Workday

```{r}
WDAY = pdfetch_YAHOO("WDAY", fields = c("open", "high", "low", "close"),
                     from = as.Date("2019-09-01"),interval = "1d")
colnames(WDAY)=c("Open","High","Low","Close")

w =tail(WDAY, n = 180)



dygraph(w,
        ylab = "Stock Price ($)",
        group = "stockprices") %>%
dyCandlestick() %>%
dyLegend(width = 400) %>%
dyRangeSelector(dateWindow = c("2019-10-03", "2020-03-09"))
```


-----------------------------------------------------------------------

### SAP

```{r}
SAP = pdfetch_YAHOO("SAP", fields = c("open", "high", "low", "close"),
                     from = as.Date("2019-09-01"),interval = "1d")
colnames(SAP)=c("Open","High","Low","Close")

s =tail(SAP, n = 180)


dygraph(s,
        ylab = "Stock Price ($)",
        group = "stockprices") %>%
dyCandlestick() %>%
dyLegend(width = 400) %>%
dyRangeSelector(dateWindow = c("2019-10-03", "2020-03-09"))
```

### Oracle

```{r}
ORCL = pdfetch_YAHOO("ORCL", fields = c("open", "high", "low", "close"),
                     from = as.Date("2019-09-01"),interval = "1d")
colnames(ORCL)=c("Open","High","Low","Close")

o =tail(ORCL, n = 180)


dygraph(o,
        ylab = "Stock Price ($)",
        group = "stockprices") %>%
dyCandlestick() %>%
dyLegend(width = 400) %>%
dyRangeSelector(dateWindow = c("2019-10-03", "2020-03-09"))

```

### Tesla

```{r}
TSLA = pdfetch_YAHOO("TSLA", fields = c("open", "high", "low", "close"),
                     from = as.Date("2019-09-01"),interval = "1d")
colnames(TSLA)=c("Open","High","Low","Close")

t =tail(TSLA, n = 180)


dygraph(t,
        ylab = "Stock Price ($)",
        group = "stockprices") %>%
dyCandlestick() %>%
dyLegend(width = 400) %>%
dyRangeSelector(dateWindow = c("2019-10-03", "2020-03-09"))

```

Analysis of stocks
=====================================

Sidebar {.sidebar}
-----------------------------------------------------------------------

### Analysis

After studying the four stocks, I believe it would be a good investment to buy Tesla stocks. The reason being that Tesla has been at a constant uptrend. Recently we are seeing some price dip but it is holding that support price and is showing signs of an uptrend. So it would be a good short term investment to buy at a low price and sell at a high price.  

Column {.tabset data-height=550}
-----------------------------------------------------------------------

### Closing Price
```{r}
fdata = pdfetch_YAHOO(identifiers, fields = c("open","adjclose"),
                      from = as.Date("2019-09-03"), to = as.Date("2020-03-09"),interval = "1d")

AdjClose = cbind(fdata$WDAY.adjclose,fdata$SAP.adjclose, fdata$ORCL.adjclose, fdata$TSLA.adjclose)
colnames(AdjClose) <-c("WDAY","SAP","ORCL","TSLA")

ac = tail(AdjClose, n = 180)


dygraph(ac,
        main = "Adjusted Closing Price (for past 180 trading days)", 
        ylab = "Closing Price ($)",
        group = "trading") %>%
dyOptions(strokeWidth = 3,colors = RColorBrewer::brewer.pal(8, "Set2"),connectSeparatedPoints = TRUE)%>%
dyHighlight(highlightCircleSize = 4, 
              highlightSeriesBackgroundAlpha = 0.2,
              hideOnMouseOut = TRUE)%>%
dyLegend(width = 400) %>%
dyRangeSelector(dateWindow = c("2019-09-03", "2020-03-09"))
```

### Volume
```{r}
vol = pdfetch_YAHOO(identifiers, fields =c("volume","adjclose"),
                    from = as.Date("2019-09-03"),interval = "1d")

vol1 = cbind(vol$WDAY.volume,vol$SAP.volume,vol$ORCL.volume,vol$TSLA.volume)
colnames(vol1) = c("WDAY","SAP","ORCL","TSLA")

v = tail(vol1, n = 180)



dygraph(v,
        main = "Trading Volume (for past 180 trading days)", 
        ylab = "Volume",
        group = "trading") %>%
dyOptions(strokeWidth = 2,colors = RColorBrewer::brewer.pal(8,"Set2"),connectSeparatedPoints = TRUE)%>%
dyHighlight(highlightCircleSize = 3, 
              highlightSeriesBackgroundAlpha = 0.2,
              hideOnMouseOut = TRUE)%>%
dyLegend(width = 400) %>%
dyRangeSelector(dateWindow = c("2019-09-03", "2020-03-09"))

```
